---
# title: "1. Model Execution"
# author: "Ignasi Puig-de-Dou, Xavier Puig, Bernat Plandolit"
# date: "24/11/2023"
output:
  github_document: default
editor_options:
  chunk_output_type: console
---

# Model Execution

This markdown file provides the code required to run the article's model.

For confidentiality reasons only the 4 printers graphed in the article are provided. 
They are printer ID's 7, 116, 207 and 216. Due to the scarce number of observations 
used (4 printers with a total of 96 observed printer-weeks), the results cannot 
reflect the actual results obtained in the article. It is intended to show the 
JAGS sampling execution. The results from the complete dataset are stored in the [model.RData](https://doi.org/10.5281/zenodo.10413762) including the 357 printers 
and 8,113 printer weeks.

## 1. Library loading

```{r libs, results='hide', message=FALSE}

library(R2jags)

```


## 2. Data loading

Only 4 out of the 357 printers are included in this code to keep data confidentiality.
Each row in the dataset is a printer-week. The dataset cotains the following fields 
(columns):

 1. **rowId**, the consecutive row number from the original 8,113 printer-weeks dataset for all 357 printers.
 2. **id**, printer ID
 3. **jagsId**, the ID needed for the JAGS model to run (01ModelExecution.Rmd) as it expects consecutively numbered
    printers. It goes from 1 to 4.
 4. **week**, observed week ISO number.
 5. **nTotal**, total number of errors logged by the printer on this week.
 6. **timeImHour**, total weekly printing time.
 7. **PVCsecPerc**, total percent time spent printing on PVC on this week.
 8. **jobsPerTime**, average number of jobs per printing hour.

```{r}

load(file='data/data.RData')

nObs = dim(data)[1]
nMaq = length(unique(data$id))

```

## 3. Bayes Model Implementation

### 3.1 Model 

$$\displaylines{
y_{jt} \sim Pois(\lambda_{jt} T_{jt}) = Pois(\theta_{jt}) \\
log(\theta_{jt}) = \beta_0 + \beta_1 \times PVCsecPerc_{jt} + \beta_2 \times jobsPerTime_{jt} + \alpha_j + z_{jt} \delta + log(T_{jt}) \\
\beta_k \sim N(0,100) \\
z_{jt} \sim Bernoulli(\pi) \\
\pi \sim Beta(1,3) \\
\alpha_j \sim N(0,\sigma^2) \\
\delta \sim Gamma(0.001, 0.001) \\
1/\sigma^2 = \tau^2 \sim Gamma(0.001,0.001) 
}$$

```{r model, eval = TRUE}

modBugs2 <- "
	model {
    
    # likelihood	
	
  	for (i in 1:nObs){
  		y[i] ~ dpois(theta[i])
  		log(theta[i]) <- b0 + b1 * pvc[i] + b2 * jobRate[i] + alpha[printer[i]] + z[i] * delta +log(offset[i])
  		z[i] ~ dbern(p)
  	}	

  	# priors
  	
  	for (j in 1:nPrinter) {
  	  alpha[j] ~ dnorm(0,tau2)
  	}
  	delta ~ dgamma(0.001,0.001)
  	tau2 ~ dgamma(0.001,0.001)
  	b0 ~ dnorm(0,1/100)
  	b1 ~ dnorm(0,1/100)
  	b2 ~ dnorm(0,1/100)
  	p ~ dbeta (1,3)

  }
"

```

### 3.2 Model Data

```{r modeldata, eval = TRUE}

dataBugs <- list(nObs = nObs,
                 nPrinter = nMaq,
                 y = data$nTotal,
                 offset = data$timeImHour,
                 printer = data$jagsId,
                 pvc = data$PVCsecPerc,
                 jobRate = data$jobsPerTime)

```

### 3.3 Model MCMC Sampling

```{r model1Exec, eval = TRUE}

Iter <- 10000

Burn <- 1000
Chain <- 3
Nthin <- 1

jag.inits <- function(){
  list('b0' = rnorm(1,-3,0.5), 'b1' = 0, 'b2' = 0, 'alpha' = rnorm(nMaq,0,1), 'delta' = runif(1, 0, 4), 
       'z' = rep(0,nObs), tau2 = 1)
}

set.seed(1964)
initPar <- lapply(1:Chain, function(i) jag.inits())

parameters <- c("b0","b1","b2","alpha","delta","p","z","tau2") 

model <- jags(dataBugs,
              inits = initPar,
              parameters.to.save=parameters,
              n.iter = (Iter+Burn),
              n.burnin = Burn,
              n.thin = Nthin,
              n.chains = Chain,
              model = textConnection(modBugs2))

```

