---
# title: "4. OC curves construction"
# author: "Ignasi Puig-de-Dou, Xavier Puig, Bernat Plandolit"
# date: "24/11/2023"
output:
  github_document: default
editor_options:
  chunk_output_type: console
---

# OC Curves Construction

This script constructs the OC curves using different print times and deltas 
and the three types of printer usage (low, medium and high). It loads the results
from the performance scenario simulation stored at data/results.RData, computes 
the required metrics for the OC curves and plots them. 

## 1. Library loading

```{r libs, results='hide', message=FALSE}

library(ggplot2)
library(dplyr)

# color blindness pallette
cbp2 <- c("#000000", "#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

```

## 2. Data loading

Loads the performancResults dataframe from the model performance simulation 
(03_ModelPerformance.Rmd). Each row in the dataframe is a combination of scenario
(9 available) and delta value (from 0 to 6 in steps of 0.25). It includes:

 1. **scenario**, scenario number, from 1 to 9 as three different usage conditions are combined with three different print times.
 2. **delta**, out of control multiplier used in the current simulation.
 3. **tp**, true positives observed in the simulation.
 4. **fn**, false negatives observed in the simulation. The sum of true positives and false negatives gives the total 1,785,000
    printer-weeks sampled per simulation (357 printers, 5,000 times each).
 5. **type**, scenario usage conditions. It is a combination of PVC printed and jobs per hour done. It has three levels: low (0% PVC and 1 job every 2 hours), medium (50% PVC and 1 job change per hour) and high (100% PVC and 3 job changes per hour).
 6.   **timeImHour**, total weekly printing time.
 7. **PVCsecPerc**, total percent time spent printing on PVC on this week.
 8. **jobsPerTime**, average number of jobs per printing hour.

```{r}

# results dataframe from the 03_ModelPerformance.Rmd script

load(file = 'data/results.RData')

```

## 3. OC curves generation

In order to plot the OC curves the script computes the beta value (false negatives).

```{r}

# compute required intermediat values

N <- results$fn[1]+results$tp[1]
results$beta <- results$fn/N
results$power <- 1-results$beta

# Plot results

ggplot(results, aes(x = delta, y = beta, col = as.factor(timeImHour), lty = as.factor(timeImHour)))+
  geom_line(linewidth = 1) +
  facet_grid( ~ type) +
  scale_y_continuous(breaks = seq(0,1,0.1)) +
  labs(color = 'printing time (h)', linetype = 'printing time (h)', y = 'false negatives', x = expression(paste(delta[0]))) +
  scale_fill_manual(values = cbp2[-1], aesthetics = c('color','fill')) 

# error type I estimation

delta0 <- results[results$delta == 0,]
colnames(delta0)[9:10] <- c('1-alpha','alpha')

delta0$alpha <- delta0$alpha*100
delta0$type <- ordered(delta0$type, levels = c('low','medium','high'))

delta0[, c('scenario','delta','alpha','type','timeImHour','PVCsecPerc','jobsPerTime')]

```

